---
title: "Project"
author: "Blake Freeman"
date: "4/15/2020"
output: html_document
---
  In a effort to create the best model for predicting Attrition and Salary, I used one hot encoding, corrlation matrix's, KNN and linear regression. The code below shows my thought process and what I did to get these results.  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
```

```{r}
Attrition_df <- read.csv("C:/Users/blake/Desktop/Case Study 2/Case Study 2/CaseStudy2-data.csv", header = TRUE)
head(Attrition_df)

Attrition_df_test <- read.csv("C:/Users/blake/Desktop/Case Study 2/Case Study 2/CaseStudy2CompSet No Attrition.csv", header = TRUE)

Salary_df_test <- read.csv("C:/Users/blake/Desktop/Case Study 2/Case Study 2/CaseStudy2CompSet_No_Salary.csv", header = TRUE)

head(Attrition_df)
head(Attrition_df_test)
```

## Including Plots

You can also embed plots, for example:

```{r}
library(ggplot2)
ggplot(Attrition_df, aes(x=Age, fill= Attrition)) + geom_histogram() + facet_wrap(~Attrition)
```
```{r}
ggplot(Attrition_df, aes(x = Gender, y = Age))+
  geom_bar(
    aes(fill = BusinessTravel), stat = "identity", color = "white",
    position = position_dodge(0.9)
    )+
  facet_wrap(~Attrition)

```

```{r}
ggplot(Attrition_df, aes(x=JobLevel, fill= Attrition)) + geom_histogram()
```

```{r}
ggplot(Attrition_df, aes(x=JobInvolvement, fill= Attrition)) + geom_histogram()
```


```{r}
ggplot(Attrition_df,aes(MaritalStatus, fill= Attrition)) + geom_bar()+ facet_wrap(~Attrition)
```

```{r}
ggplot(Attrition_df,aes(Department, fill= Attrition)) + geom_bar()+ facet_wrap(~Attrition)
```


```{r}
ggplot(Attrition_df,aes(Over18, fill= Attrition)) + geom_bar()
```

#Encoded all the values to be able to get a more clear picture when doing a analyisis. 
#Also turning Attrition into a bollien value where 1 is true for Attrition. 
```{r}
Attrition_df_he <- Attrition_df

for(unique_value in unique(Attrition_df_he$BusinessTravel)){
 

Attrition_df_he[paste("BusinessTravel", unique_value, sep = ".")] <- ifelse(Attrition_df_he$BusinessTravel == unique_value, 1, 0)

}

for(unique_value in unique(Attrition_df_he$Department)){
 

Attrition_df_he[paste("Department", unique_value, sep = ".")] <- ifelse(Attrition_df_he$Department == unique_value, 1, 0)

}


for(unique_value in unique(Attrition_df_he$EducationField)){
 

Attrition_df_he[paste("EducationField", unique_value, sep = ".")] <- ifelse(Attrition_df_he$EducationField == unique_value, 1, 0)

}


for(unique_value in unique(Attrition_df_he$Gender)){
 

Attrition_df_he[paste("Gender", unique_value, sep = ".")] <- ifelse(Attrition_df_he$Gender == unique_value, 1, 0)

}

for(unique_value in unique(Attrition_df_he$JobRole)){
 

Attrition_df_he[paste("JobRole", unique_value, sep = ".")] <- ifelse(Attrition_df_he$JobRole == unique_value, 1, 0)

}

for(unique_value in unique(Attrition_df_he$Over18)){
 

Attrition_df_he[paste("Over18", unique_value, sep = ".")] <- ifelse(Attrition_df_he$Over18 == unique_value, 1, 0)

}


for(unique_value in unique(Attrition_df_he$OverTime)){
 

Attrition_df_he[paste("OverTime", unique_value, sep = ".")] <- ifelse(Attrition_df_he$OverTime == unique_value, 1, 0)

}

for(unique_value in unique(Attrition_df_he$Attrition)){
 

Attrition_df_he[paste("Attrition", unique_value, sep = ".")] <- ifelse(Attrition_df_he$Attrition == unique_value, 1, 0)

}


```
#One hot encoding of the Attrtion Test
```{r}
Attrition_df_test_he <- Attrition_df_test

for(unique_value in unique(Attrition_df_test_he$BusinessTravel)){
 

Attrition_df_test_he[paste("BusinessTravel", unique_value, sep = ".")] <- ifelse(Attrition_df_test_he$BusinessTravel == unique_value, 1, 0)

}

for(unique_value in unique(Attrition_df_test_he$Department)){
 

Attrition_df_test_he[paste("Department", unique_value, sep = ".")] <- ifelse(Attrition_df_test_he$Department == unique_value, 1, 0)

}


for(unique_value in unique(Attrition_df_test_he$EducationField)){
 

Attrition_df_test_he[paste("EducationField", unique_value, sep = ".")] <- ifelse(Attrition_df_test_he$EducationField == unique_value, 1, 0)

}


for(unique_value in unique(Attrition_df_test_he$Gender)){
 

Attrition_df_test_he[paste("Gender", unique_value, sep = ".")] <- ifelse(Attrition_df_test_he$Gender == unique_value, 1, 0)

}

for(unique_value in unique(Attrition_df_test_he$JobRole)){
 

Attrition_df_test_he[paste("JobRole", unique_value, sep = ".")] <- ifelse(Attrition_df_test_he$JobRole == unique_value, 1, 0)

}

for(unique_value in unique(Attrition_df_test_he$Over18)){
 

Attrition_df_test_he[paste("Over18", unique_value, sep = ".")] <- ifelse(Attrition_df_test_he$Over18 == unique_value, 1, 0)

}


for(unique_value in unique(Attrition_df_test_he$OverTime)){
 

Attrition_df_test_he[paste("OverTime", unique_value, sep = ".")] <- ifelse(Attrition_df_test_he$OverTime == unique_value, 1, 0)

}

for(unique_value in unique(Attrition_df_test_he$Attrition)){
 

Attrition_df_test_he[paste("Attrition", unique_value, sep = ".")] <- ifelse(Attrition_df_test_he$Attrition == unique_value, 1, 0)

}

head(Attrition_df_test_he)
```



```{r}
Attrition_df_he.m <- as.matrix(Attrition_df_he)

```

```{r}
is.na(Attrition_df_he)
apply(is.na(Attrition_df_he), 2, which)
```
#droping the variables I hot encoded and then any boolen values. 
```{r}
drops <- c("Attrition","BusinessTravel","Department","EducationField","Gender","JobRole","Over18","OverTime", "MaritalStatus","ID","EmployeeCount","Over18.Y","StandardHours","Attrition.No","OverTime.Yes","Gender.Female")

Attrition_df_he <- Attrition_df_he[ , !(names(Attrition_df_he) %in% drops)]


head(Attrition_df_he)

```




```{r}

library("Hmisc")
Attrition_df_he.corr <- rcorr(as.matrix(Attrition_df_he))

```



```{r}

Attrition_df_he.P <- as.data.frame(t(Attrition_df_he.corr$P))

Attrition_df_he.P

```
Age 
DistanceFromHome
JobInvolvement
JobLevel
JobSatisfaction
MonthlyIncome
StockOptionLevel
TotalWorkingYears
WorkLifeBalance
YearsAtCompany
YearsInCurrentRole
Department.Sales
Department.Research & Development
JobRole.Research Director
JobRole.Manufacturing Director
JobRole.Sales Representaitive
OverTime.No
# Corrlation Matrix
```{r,fig.width=35,fig.height=20}
library(corrplot)
corrplot(Attrition_df_he.corr$r, type="upper",  
         p.mat = Attrition_df_he.corr$P, sig.level = 0.05, insig = "blank")
```

```{r}
Attrition_df_he.corr_df <- as.data.frame(t(Attrition_df_he.corr$P))
Attrition_df_he.corr_df

write.csv(Attrition_df_he.corr_df,"C:/Users/blake/Desktop/Case Study 2/Case Study 2/Attrition_corr.csv", row.names = FALSE)
```


#Variables I am keeping for Attrition
```{r}
Keep <- c("Attrition.Yes",
"Age", 
"DistanceFromHome",
"JobInvolvement",
"JobLevel",
"JobSatisfaction",
"MonthlyIncome",
"StockOptionLevel",
"TotalWorkingYears",
"WorkLifeBalance",
"YearsAtCompany",
"YearsInCurrentRole",
"Department.Sales",
"Department.Research & Development",
"JobRole.Research Director",
"JobRole.Manufacturing Director",
"JobRole.Sales Representaitive",
"OverTime.No"
)


Attrition_df_he <- Attrition_df_he[ , (names(Attrition_df_he) %in% Keep)]
Attrition_df_he
```

#Variables I am keeping for Attrition 
```{r}
Keep <- c("Attrition.Yes",
"Age", 
"DistanceFromHome",
"JobInvolvement",
"JobLevel",
"JobSatisfaction",
"MonthlyIncome",
"StockOptionLevel",
"TotalWorkingYears",
"WorkLifeBalance",
"YearsAtCompany",
"YearsInCurrentRole",
"Department.Sales",
"Department.Research & Development",
"JobRole.Research Director",
"JobRole.Manufacturing Director",
"JobRole.Sales Representaitive",
"OverTime.No"
)


Attrition_df_test_he <- Attrition_df_test_he[ , (names(Attrition_df_test_he) %in% Keep)]
Attrition_df_test_he
```


#Corrlation Matrix for Attrition
```{r width=35,fig.height=20}
Attrition_df_he.corr2 <- rcorr(as.matrix(Attrition_df_he))


corrplot(Attrition_df_he.corr2$r, type="upper",  
         p.mat = Attrition_df_he.corr2$P, sig.level = 0.05, insig = "blank")
```

```{r}
Attrition_df_he.corr2
```


#KNN to test best accuarcy 
```{r}
library(tidyverse)
library(class)
library(caret)
library(e1071)
library(ggplot2)
```
```{r}
set.seed(500)
iterations = 100
numks = 60
splitPerc = .70

masterAcc = matrix(nrow = iterations, ncol = numks)

for(j in 1:iterations)
{
   trainIndices = sample(1:dim(Attrition_df_he)[1],round(splitPerc * dim(Attrition_df_he)[1]))
    train = Attrition_df_he[trainIndices,]
    test = Attrition_df_he[-trainIndices,]
  for(i in 1:numks)
  {
    classifications = knn(train[,c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16)],test[,c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16)],train$Attrition.Yes, prob = TRUE, k = i)
    u <- union(classifications,test$Attrition.Yes)
    t <- table(factor(classifications, u), factor(test$Attrition.Yes, u))
    CM = confusionMatrix(t)
    masterAcc[j,i] = CM$overall[1]
  }
  
}

MeanAcc = colMeans(masterAcc)

plot(seq(1,numks,1),MeanAcc, type = "l")

which.max(MeanAcc)
max(MeanAcc)
```


#KNN model for Attrition
```{r}
Attrition_pred <- knn(train[,c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16)],Attrition_df_test_he[,c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16)],cl= train$Attrition.Yes, k = 36)
Attrition_pred_df <- as.data.frame(Attrition_pred)
Attrition_pred_df

```
```{r}

Attrition_pred_df$Attrition <- "No"
Attrition_pred_df$Attrition[Attrition_pred_df$Attrition_pred >1] <- "Yes"
Attrition_pred_df
```
#One Hot accoding for Salary
```{r}
Salary <- Attrition_df

for(unique_value in unique(Salary$BusinessTravel)){
 

Salary[paste("BusinessTravel", unique_value, sep = ".")] <- ifelse(Salary$BusinessTravel == unique_value, 1, 0)

}

for(unique_value in unique(Salary$Department)){
 

Salary[paste("Department", unique_value, sep = ".")] <- ifelse(Salary$Department == unique_value, 1, 0)

}


for(unique_value in unique(Salary$EducationField)){
 

Salary[paste("EducationField", unique_value, sep = ".")] <- ifelse(Salary$EducationField == unique_value, 1, 0)

}


for(unique_value in unique(Salary$Gender)){
 

Salary[paste("Gender", unique_value, sep = ".")] <- ifelse(Salary$Gender == unique_value, 1, 0)

}

for(unique_value in unique(Salary$JobRole)){
 

Salary[paste("JobRole", unique_value, sep = ".")] <- ifelse(Salary$JobRole == unique_value, 1, 0)

}




for(unique_value in unique(Salary$OverTime)){
 

Salary[paste("OverTime", unique_value, sep = ".")] <- ifelse(Salary$OverTime == unique_value, 1, 0)

}

for(unique_value in unique(Salary$Attrition)){
 

Salary[paste("Attrition", unique_value, sep = ".")] <- ifelse(Salary$Attrition == unique_value, 1, 0)

}


drops <- c("Attrition","BusinessTravel","Department","EducationField","Gender","JobRole","Over18","OverTime", "MaritalStatus","ID","EmployeeCount","Over18.Y","StandardHours","Attrition.No","OverTime.Yes","Gender.Female")

Salary <- Salary[ , !(names(Salary) %in% drops)]


head(Salary)

```
#Corrlation Matrix for Salary
```{r}
Salary.corr <- rcorr(as.matrix(Salary))

Salary.corr.P <- as.data.frame(t(Salary.corr$P))

Salary.corr.P 

```
"TotalWorkingYears",
"YearsAtCompany",
"YearsInCurrentRole",
"YearsSinceLastPromotion",
"YearsWithCurrManager",
"JobRole.Research Director",
"JobRole.Manufacturing Director",
"JobRole.Research Scientist",
"JobRole.Sales Representative",
"JobRole.Healthcare Representative",
"JobRole.Manager",
"JobRole.Human Resources",
"JobRole.Laboratory Technician",
"Attrition.Yes",

#Variables I am keeping
```{r width=35,fig.height=20}
Keep <- c("TotalWorkingYears",
"YearsAtCompany",
"YearsWithCurrManager",
"JobRole.Research Director",
"JobRole.Manufacturing Director",
"JobRole.Research Scientist",
"JobRole.Sales Representative",
"JobRole.Manager",
"JobRole.Human Resources",
"JobRole.Laboratory Technician",
"MonthlyIncome"

)


Salary <- Salary[ , (names(Salary) %in% Keep)]

Salary.corr <- rcorr(as.matrix(Salary))


corrplot(Salary.corr$r, type="upper",  
         p.mat = Salary$P, sig.level = 0.05, insig = "blank")
```


#one Hot encoding Salary df
```{r}
Salary_test <- Salary_df_test

for(unique_value in unique(Salary_test$BusinessTravel)){
 

Salary_test[paste("BusinessTravel", unique_value, sep = ".")] <- ifelse(Salary_test$BusinessTravel == unique_value, 1, 0)

}

for(unique_value in unique(Salary_test$Department)){
 

Salary_test[paste("Department", unique_value, sep = ".")] <- ifelse(Salary_test$Department == unique_value, 1, 0)

}

for(unique_value in unique(Salary_test$EducationField)){
 

Salary_test[paste("EducationField", unique_value, sep = ".")] <- ifelse(Salary_test$EducationField == unique_value, 1, 0)

}


for(unique_value in unique(Salary_test$Gender)){
 

Salary_test[paste("Gender", unique_value, sep = ".")] <- ifelse(Salary_test$Gender == unique_value, 1, 0)

}

for(unique_value in unique(Salary_test$JobRole)){
 

Salary_test[paste("JobRole", unique_value, sep = ".")] <- ifelse(Salary_test$JobRole == unique_value, 1, 0)

}




for(unique_value in unique(Salary_test$OverTime)){
 

Salary_test[paste("OverTime", unique_value, sep = ".")] <- ifelse(Salary_test$OverTime == unique_value, 1, 0)

}

for(unique_value in unique(Salary_test$Attrition)){
 

Salary_test[paste("Attrition", unique_value, sep = ".")] <- ifelse(Salary_test$Attrition == unique_value, 1, 0)

}

Keep <- c("TotalWorkingYears",
"YearsAtCompany",


"YearsWithCurrManager",
"JobRole.Research Director",
"JobRole.Manufacturing Director",
"JobRole.Research Scientist",
"JobRole.Sales Representative",

"JobRole.Manager",
"JobRole.Human Resources",
"JobRole.Laboratory Technician"

)

Salary_test <- Salary_test[ , (names(Salary_test) %in% Keep)]


head(Salary_test)
```
```{r}
Salary_df_test
```
#Linear Regression Model for Salary
```{r}
library(dvmisc)
library(Metrics)

set.seed(100)

train <- createDataPartition(Salary$MonthlyIncome, p=0.95, list=FALSE)
training <- Salary[ train, ]
test <- Salary[-train,]


Model1_fit = lm(MonthlyIncome ~ ., data = training)

sm <- summary(Model1_fit)


MSE <- get_mse(Model1_fit, var.estimate = FALSE)


predictions_model = predict(Model1_fit,test)
predictions = predict(Model1_fit,Salary_test)


predictions_df <- as.data.frame(predictions)
predictions_df
sm
rmse(training$MonthlyIncome, predictions_model)

```
#Adding the Predtions to the Test Sets for output. 
```{r}
Attrition_predtions <- cbind(Attrition_df_test,Attrition_pred_df$Attrition)

Salary_predtions <- cbind(Salary_df_test,predictions_df$predictions)

names(Attrition_predtions)[names(Attrition_predtions) == "Attrition_pred_df$Attrition"] <- "Predicted Attrition"
names(Salary_predtions)[names(Salary_predtions) == "predictions_df$predictions"] <- "Predicted Monthly Income"


head(Attrition_predtions)
head(Salary_predtions)


write.csv(Attrition_predtions,"C:/Users/blake/Desktop/Case Study 2/CaseStudy2DDS-/Case2Predictions Blake Freeman Attrition.csv", row.names = FALSE)

write.csv(Salary_predtions,"C:/Users/blake/Desktop/Case Study 2/CaseStudy2DDS-/Case2Predictions Blake Freeman Salary.csv", row.names = FALSE)
```

